Decomposing labelled proof theory for intuitionistic modal logic

### 2019-09-06 caleidoscope/caleidoscope.tex
small modif

We introduce the rules that represent the semantics of modal operators $\square$ and $\Diamond$. The system $\labIKp$ presents the next four rules (two for $\square$ and two for $\Diamond$): 
\begin{center}
	\begin{minipage}{.95\textwidth}
		\begin{tabular}{@{\!}c@{\quad}c}
			\\\\
			$\vlinf{\llabrn\BOX}{}{\B, x \le y, yRz, \Left, \labels{x}{\BOX A} \SEQ \Right}{\B, x \le y, yRz, \Left, \labels{x}{\BOX A}, \labels{z}{A} \SEQ \Right}$
			&
			$\vlinf{\rlabrn\BOX}{\text{\scriptsize $y, z$ fresh}}{\B, \Left \SEQ \Right, \labels{x}{\BOX A}}{\B, x \le y, y \rel z, \Left \SEQ \Right, \labels{z}{A}}$
			\\\\
			$\vlinf{\llabrn\DIA}{\text{\scriptsize $y$ fresh}}{\B, \Left, \labels{x}{\DIA A} \SEQ \Right}{\B, x \rel y, \Left, \labels{y}{A} \SEQ \Right}$
			&
			$\vlinf{\rlabrn\DIA}{}{\B, x \rel y, \Left \SEQ \Right, \labels{x}{\DIA A}}{\B, x \rel y, \Left \SEQ \Right, \labels{x}{\DIA A}, \labels{y}{A}}$
			\\
		\end{tabular}
	\end{minipage}
\end{center}

@@

\begin{itemize}
	\item Rule $\llabrn\BOX$.
	Continue with the same reasoning, $\B$ is a set of relational and preorder atoms and $\Left$ and $\Right$ are multiset of laballed formulas. Semantically, the rule expresses, from the conclusion to the premise, that given $x, y$ and $z$ such that $x \le y$ and $yRz$ where the formula $\square A$ is true in $x$, then the formula $A$ is true in $z$.

	\item Rule $\rlabrn\DIA$.
	Semantically, the diamond establishes that in a model $\M$ and a world $v$ the formula $\Diamond A$ is satisfied, if and only if, there exists a world $u$ such that $vRu$ and the formula $A$ is satisfied in $u$. In order to capture this definition, the rule shows that if $xRy$ and the formula $\Diamond A$ is satisfied in $x$, then the formula $A$ is true in $y$.

@@

\begin{proof}
	\begin{center}
		\scalebox{0.93}{
			$\vlderivation {\vlin {\sir}
				{y$ fresh$}
				{\Rightarrow x \colon \lozenge \square A \vljm \square \lozenge A}
				{\vlin {\sbr}
					{z, w $ fresh $}
					{x \le y, y \colon \lozenge \square A \Rightarrow y \colon \square \lozenge A}
					{\vlin {\sdl}
						{u$ fresh$}
						{x \le y, y \le z, zRw, y \colon \lozenge \square A \Rightarrow w \colon \lozenge A}
						{\vlin {\ftwo}
							{t$ fresh$}
							{x \le y, y \le z, zRw, yRu, u \colon \square A \Rightarrow w \colon \lozenge A}
							{\vlin {\gklmn}
								{t', j$ fresh$}
								{x \le y, y \le z, u \le t, zRw, yRu, zRt, u \colon \square A \Rightarrow w \colon \lozenge A}
								{\vlin {\sdr}
									{}
									{x \le y, y \le z, u \le t, t \le t', zRw, yRu, zRt, t'Rj, wRj, u \colon \square A \Rightarrow w \colon \lozenge A}
									{\vlin {\trans}
										{}
										{x \le y, y \le z, u \le t, t \le t', zRw, yRu, zRt, t'Rj, wRj, u \colon \square A \Rightarrow w \colon \lozenge A, j \colon A}
										{\vlin {\sbl}
											{}
											{x \le y, y \le z, u \le t, t \le t', u \le t', zRw, yRu, zRt, t'Rj, wRj, u \colon \square A \Rightarrow w \colon \lozenge A, j \colon A}
											{\vlin {\refl}
												{}
												{x \le y, y \le z, u \le t, t \le t', u \le t', zRw, yRu, zRt, t'Rj, wRj, u \colon \square A, j \colon A \Rightarrow w \colon \lozenge A, j \colon A}
												{\vlin {\ids}
													{}
													{x \le y, y \le z, u \le t, t \le t', u \le t',j \le j, zRw, yRu, zRt, t'Rj, wRj, u \colon \square A, j \colon A \Rightarrow w \colon \lozenge A, j \colon A}
													{\vlhy {}}}}}}}}}}}}$}
	\end{center}
\end{proof}

@@

\begin{lemma}\label{lemma:admis} The following rules are admissible in $\labIKp$:
	\begin{enumerate}
		\item{$\vlderivation {\vlin {\boxlk}{}{\B, \Left, x(\le \circ $R$)^{k}y, x \colon \square^{k} A\Rightarrow \Right}{\vlhy {\B, \Left, x(\le \circ $R$)^{k}y, x \colon \square^{k} A, z \colon A \Rightarrow \Right}}}$}
		\item{$\vlderivation{\vlin {\boxk}{}{\B, \Left \Rightarrow \Right, x \colon \square^{k} A}{\vlhy {\B, x(\le \circ $R$)^{k}y,\Left \Rightarrow \Right, y \colon A}}}$}
		\item{$\vlderivation { \vlin {\diamk}{}{\B, \Left, x \colon \lozenge^{k} A \Rightarrow \Right}{\vlhy {\B, xR^{k}y, \Left, y \colon A \Rightarrow \Right}}}$ }
		\item{$\vlderivation { \vlin {\diamrk}{}{\B, \Left, xR^{k}y \Rightarrow \Right, x \colon \lozenge^{k}A}{\vlhy {\B, \Left, xR^{k}y \Rightarrow \Right, x \colon \lozenge^{k}A, y \colon A}}}$}
	\end{enumerate}
	%\textbf{Probably F2g???}
\end{lemma}

>>>>

We introduce the rules that represent the semantics of modal operators $\BOX$ and $\DIA$. The system $\labIKp$ presents the next four rules (two for $\BOX$ and two for $\DIA$): 
\begin{center}
	\begin{minipage}{.95\textwidth}
		\begin{tabular}{@{\!}c@{\quad}c}
			\\\\
			$\vlinf{\llabrn\BOX}{}{\B, x \le y, yRz, \Left, \labels{x}{\BOX A} \SEQ \Right}{\B, x \le y, yRz, \Left, \labels{x}{\BOX A}, \labels{z}{A} \SEQ \Right}$
			&
			$\vlinf{\rlabrn\BOX}{\text{\scriptsize $y, z$ fresh}}{\B, \Left \SEQ \Right, \labels{x}{\BOX A}}{\B, x \le y, y \rel z, \Left \SEQ \Right, \labels{z}{A}}$
			\\\\
			$\vlinf{\llabrn\DIA}{\text{\scriptsize $y$ fresh}}{\B, \Left, \labels{x}{\DIA A} \SEQ \Right}{\B, x \rel y, \Left, \labels{y}{A} \SEQ \Right}$
			&
			$\vlinf{\rlabrn\DIA}{}{\B, x \rel y, \Left \SEQ \Right, \labels{x}{\DIA A}}{\B, x \rel y, \Left \SEQ \Right, \labels{x}{\DIA A}, \labels{y}{A}}$
			\\
		\end{tabular}
	\end{minipage}
\end{center}

@@

\begin{itemize}
	\item Rule $\llabrn\BOX$.
	Continue with the same reasoning, $\B$ is a set of relational and preorder atoms and $\Left$ and $\Right$ are multiset of laballed formulas. Semantically, the rule expresses, from the conclusion to the premise, that given $x, y$ and $z$ such that $x \le y$ and $yRz$ where the formula $\BOX A$ is true in $x$, then the formula $A$ is true in $z$.

	\item Rule $\rlabrn\DIA$.
	Semantically, the diamond establishes that in a model $\M$ and a world $v$ the formula $\DIA A$ is satisfied, if and only if, there exists a world $u$ such that $vRu$ and the formula $A$ is satisfied in $u$. In order to capture this definition, the rule shows that if $xRy$ and the formula $\DIA A$ is satisfied in $x$, then the formula $A$ is true in $y$.

@@

\begin{proof}
	\begin{center}
		\scalebox{0.93}{
			$\vlderivation {\vlin {\sir}
				{y$ fresh$}
				{\Rightarrow \labels{x}{\DIA \BOX A \IMP \BOX \DIA A}}
				{\vlin {\sbr}
					{z, w $ fresh $}
					{x \le y, \labels{y}{\DIA \BOX A} \Rightarrow \labels{y}{\BOX \DIA A}}
					{\vlin {\sdl}
						{u$ fresh$}
						{x \le y, y \le z, zRw, \labels{y}{\DIA \BOX A} \Rightarrow \labels{w}{\DIA A}}
						{\vlin {\ftwo}
							{t$ fresh$}
							{x \le y, y \le z, zRw, yRu, \labels{u}{\BOX A} \Rightarrow \labels{w}{\DIA A}}
							{\vlin {\gklmn}
								{t', j$ fresh$}
								{x \le y, y \le z, u \le t, zRw, yRu, zRt, \labels{u}{\BOX A} \Rightarrow \labels{w}{\DIA A}}
								{\vlin {\sdr}
									{}
									{x \le y, y \le z, u \le t, t \le t', zRw, yRu, zRt, t'Rj, wRj, \labels{u}{\BOX A} \Rightarrow \labels{w}{\DIA A}}
									{\vlin {\trans}
										{}
										{x \le y, y \le z, u \le t, t \le t', zRw, yRu, zRt, t'Rj, wRj, \labels{u}{\BOX A} \Rightarrow \labels{w}{\DIA A}, \labels{j}{A}}
										{\vlin {\sbl}
											{}
											{x \le y, y \le z, u \le t, t \le t', u \le t', zRw, yRu, zRt, t'Rj, wRj, \labels{u}{\BOX A} \Rightarrow \labels{w}{\DIA A}, \labels{j}{A}}
											{\vlin {\refl}
												{}
												{x \le y, y \le z, u \le t, t \le t', u \le t', zRw, yRu, zRt, t'Rj, wRj, \labels{u}{\BOX A}, \labels{j}{A} \Rightarrow \labels{w}{\DIA A}, \labels{j}{A}}
												{\vlin {\ids}
													{}
													{x \le y, y \le z, u \le t, t \le t', u \le t',j \le j, zRw, yRu, zRt, t'Rj, wRj, \labels{u}{\BOX A}, \labels{j}{A} \Rightarrow \labels{w}{\DIA A}, \labels{j}{A}}
													{\vlhy {}}}}}}}}}}}}$}
	\end{center}
\end{proof}

@@

\begin{lemma}\label{lemma:admis} The following rules are admissible in $\labIKp$:
	\begin{enumerate}
		\item{$\vlderivation {\vlin {\boxlk}{}{\B, \Left, x(\le \circ $R$)^{k}y, \labels{x}{\BOX^{k} A}\Rightarrow \Right}{\vlhy {\B, \Left, x(\le \circ $R$)^{k}y, \labels{x}{\BOX^{k} A}, \labels{z}{A} \Rightarrow \Right}}}$}
		\item{$\vlderivation{\vlin {\boxk}{}{\B, \Left \Rightarrow \Right, \labels{x}{\BOX^{k} A}}{\vlhy {\B, x(\le \circ $R$)^{k}y,\Left \Rightarrow \Right, \labels{y}{A}}}}$}
		\item{$\vlderivation { \vlin {\diamk}{}{\B, \Left, \labels{x}{\DIA^{k} A} \Rightarrow \Right}{\vlhy {\B, xR^{k}y, \Left, \labels{y}{A} \Rightarrow \Right}}}$ }
		\item{$\vlderivation { \vlin {\diamrk}{}{\B, \Left, xR^{k}y \Rightarrow \Right, \labels{x}{\DIA^{k}A}}{\vlhy {\B, \Left, xR^{k}y \Rightarrow \Right, \labels{x}{\DIA^{k}A}, \labels{y}{A}}}}$}
	\end{enumerate}
	%\textbf{Probably F2g???}
\end{lemma}
